Fixing Vulnerabilities Hinders Maintainability

### 2019-02-04 paper/ICPC19/icpc19.tex
comentarios listing

There are security flaws that are more difficult to refactor than others. For
instance, implementing secure authentication is not as easy as fixing a
cross-site scripting vulnerability since usually the last one is can be fixed
without adding new lines of code, as seen in Figure~\ref{lst:fix}. The developer
added the function \emph{htmlentities} to \textit{escape} the data given by the
variable \emph{\$\_['file']}. Knowing which patterns are more likely to increase
maintainability issues is one step forward to bring awareness to security
engineers of what patterns need more attention. The taxonomy of security
patterns used to answer this question is the same as the one provided by the
authors of the dataset~\cite{Reis:2017:IJSSE}. Maintainability is measured
separately for each pattern.

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix}]
 <ul>
    <li class='error'>
       <?php echo $l->t('Cloud not found');?><br/>
-      <p class='hint'><?php if(isset($_['file'])) echo $_['file']?></p>
+      <p class='hint'><?php if(isset($_['file'])) echo htmlentities($_['file'])?></p>
    </li>
 </ul>

\end{lstlisting}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The
methodology is comprised of the following steps, as illustrated in
Figure~\ref{fig:met}:

>>>>

There are security flaws that are more difficult to refactor than others. For
instance, implementing secure authentication is not as easy as fixing a
cross-site scripting vulnerability since usually the last one is can be fixed
without adding new lines of code, as seen in Figure~\ref{lst:fix}.\Luis{as seen
in Listing?} The developer added the function \emph{htmlentities} to
\textit{escape} the data given by the variable \emph{\$\_['file']}.\Luis{usa
sempre o mesmo tipo de letra para nomes de funcoes. \\texttt?} Knowing which
patterns are more likely to increase maintainability issues is one step forward
to bring awareness to security engineers of what patterns need more attention.
The taxonomy of security patterns used to answer this question is the same as
the one provided by the authors of the dataset~\cite{Reis:2017:IJSSE}.
Maintainability is measured separately for each pattern.

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix}]
 <ul>
    <li class='error'>
       <?php echo $l->t('Cloud not found');?><br/>
-      <p class='hint'><?php if(isset($_['file'])) echo $_['file']?></p>
+      <p class='hint'><?php if(isset($_['file'])) echo htmlentities($_['file'])?><?php 
    </li>
 </ul>

\end{lstlisting} \Luis{no listing lst:fix o que achas de fazer linebreak para
cada uma das tags? assim nao havi o problema de ter a mesma linha dividida e ficava mais facil de ler:\\
\texttt{<p class='hint'>}\\
\texttt{~~<?php }\\
\texttt{</p>}\\
Also: reduz numero de espacos para cada tab -- 2 é suficiente.
}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The
methodology is comprised of the following steps, as illustrated in
Figure~\ref{fig:met}:
